### 2014-09-10 paper/paper.Rmd
trying to fix gender maps

```{r gendermath, fig.width=8, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
student2012.sub <- student2012[, c(1, 12, 501, 541, 546, 634)]
colnames(student2012.sub)[1] <- "name"
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub$PV1READ <- as.numeric(student2012.sub$PV1READ)
student2012.sub$PV1SCIE <- as.numeric(student2012.sub$PV1SCIE)
student2012.sub$SENWGT_STU <- as.numeric(student2012.sub$SENWGT_STU)
student2012.sub.summary.gap <- summarise(group_by(student2012.sub, name), 
                  mathgap=mean(PV1MATH[ST04Q01=="Male"], na.rm=T)-
                          mean(PV1MATH[ST04Q01=="Female"], na.rm=T),
                  wmathgap=weighted.mean(PV1MATH[ST04Q01=="Male"], w=SENWGT_STU[ST04Q01=="Male"], na.rm=T)-
                          weighted.mean(PV1MATH[ST04Q01=="Female"], w=SENWGT_STU[ST04Q01=="Female"], na.rm=T),
                  mtest.stat = t.test(PV1MATH[ST04Q01=="Male"], PV1MATH[ST04Q01=="Female"])$statistic, 
                  mp.value = t.test(PV1MATH[ST04Q01=="Male"], PV1MATH[ST04Q01=="Female"])$p.value,
                  readgap=mean(PV1READ[ST04Q01=="Male"], na.rm=T)-mean(PV1READ[ST04Q01=="Female"], na.rm=T),
                  rtest.stat = t.test(PV1READ[ST04Q01=="Male"], PV1READ[ST04Q01=="Female"])$statistic, 
                  rp.value = t.test(PV1READ[ST04Q01=="Male"], PV1READ[ST04Q01=="Female"])$p.value,
                  sciencegap=mean(PV1SCIE[ST04Q01=="Male"], na.rm=T)-mean(PV1SCIE[ST04Q01=="Female"], na.rm=T),
                  stest.stat = t.test(PV1SCIE[ST04Q01=="Male"], PV1SCIE[ST04Q01=="Female"])$statistic, 
                  sp.value = t.test(PV1SCIE[ST04Q01=="Male"], PV1SCIE[ST04Q01=="Female"])$p.value,                                         
                  minmale=min(PV1MATH[ST04Q01=="Male"], na.rm=T), 
                  minfemale=min(PV1MATH[ST04Q01=="Female"], na.rm=T), 
                  maxmale=max(PV1MATH[ST04Q01=="Male"], na.rm=T), 
                  maxfemale=max(PV1MATH[ST04Q01=="Female"], na.rm=T), 
                  propmale=length(PV1MATH[ST04Q01=="Male"])/length(PV1MATH), 
                  propfemale=length(PV1MATH[ST04Q01=="Female"])/length(PV1MATH))
#qplot(mathgap, wmathgap, data=student2012.sub.summary.gap, xlab="Mean", ylab="Weighted Mean",
#      xlim=c(-30,30), ylim=c(-30,30)) + geom_abline(slope=1) + theme(aspect.ratio=1)
student2012.sub.summary.gap$msig <- ifelse(student2012.sub.summary.gap$mp.value>0.05, "none", TRUE)
student2012.sub.summary.gap$msig[student2012.sub.summary.gap$msig==TRUE&student2012.sub.summary.gap$mtest.stat>0] <- "male"
student2012.sub.summary.gap$msig[student2012.sub.summary.gap$msig==TRUE&student2012.sub.summary.gap$mtest.stat<0] <- "female"
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$mathgap)])
qplot(name, mathgap, data=student2012.sub.summary.gap, size=propmale, color=msig) + 
  xlab("") +
  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  scale_y_continuous("Math Score Gap", breaks=seq(-30, 30, 5)) + 
  scale_size("Prop male") +
   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw() + theme(legend.position="bottom")
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$maxmale)])
#qplot(name, maxmale, data=student2012.sub.summary.gap, color=I("skyblue")) + 
#  xlab("") +
#  geom_point(aes(y=maxfemale), colour="pink") + 
#  scale_y_continuous("High Math") + coord_flip() + theme_bw()
#ggsave("gendermathtop.pdf", width=3.5, height=7)
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$minfemale)])
#qplot(name, minmale, data=student2012.sub.summary.gap, color=I("skyblue")) + 
#  xlab("") +
#  geom_point(aes(y=minfemale), colour="pink") + 
#  scale_y_continuous("Low Math") + coord_flip() + theme_bw()
#ggsave("gendermathbottom.pdf", width=3.5, height=7)
#
student2012.sub.summary.gap$rsig <- ifelse(student2012.sub.summary.gap$rp.value>0.05, "none", TRUE)
student2012.sub.summary.gap$rsig[student2012.sub.summary.gap$rsig==TRUE&student2012.sub.summary.gap$rtest.stat>0] <- "male"
student2012.sub.summary.gap$rsig[student2012.sub.summary.gap$rsig==TRUE&student2012.sub.summary.gap$rtest.stat<0] <- "female"
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$readgap)])
qplot(name, readgap, data=student2012.sub.summary.gap, size=propmale, color=rsig) + 
  xlab("") +
  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  scale_y_continuous("Reading Score Gap") + 
  scale_size("Prop male") +
   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw() + theme(legend.position="bottom")
#ggsave("genderreadgap.pdf", width=5, height=8)
#student2012.sub.summary.gap$ssig <- ifelse(student2012.sub.summary.gap$sp.value>0.05, "none", TRUE)
#student2012.sub.summary.gap$ssig[student2012.sub.summary.gap$ssig==TRUE&student2012.sub.summary.gap$stest.stat>0] <- "male"
#student2012.sub.summary.gap$ssig[student2012.sub.summary.gap$ssig==TRUE&student2012.sub.summary.gap$stest.stat<0] <- "female"
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$sciencegap)])
#qplot(name, sciencegap, data=student2012.sub.summary.gap, size=propmale, color=ssig) + 
#  xlab("") +
#  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
#  scale_y_continuous("Science Score Gap") + 
#  scale_size("Prop male") +
#   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw()
```

```{r gendermaps, fig.width=12, fig.height=7, warning=FALSE, message=FALSE, echo=FALSE}
student2012.sub.map <- left_join(student2012.sub.summary.gap, world.polys)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=msig)) +
  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  new_theme_empty + theme(legend.position="none")
#ggsave("gendermathmap.pdf", width=12, height=7)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=rsig)) +
  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  new_theme_empty + theme(legend.position="none")
#ggsave("genderreadmap.pdf", width=12, height=7)
#ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
#  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=ssig)) +
#  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
#  new_theme_empty + theme(legend.position="none")
```

>>>>

```{r gendermath, fig.width=8, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
student2012.sub <- student2012[, c(1, 12, 501, 541, 546, 634)]
student2012.sub.summary.gap <- summarise(group_by(student2012.sub, name), 
                  mathgap=mean(PV1MATH[ST04Q01=="Male"], na.rm=T)-
                          mean(PV1MATH[ST04Q01=="Female"], na.rm=T),
                  wmathgap=weighted.mean(PV1MATH[ST04Q01=="Male"], w=SENWGT_STU[ST04Q01=="Male"], na.rm=T)-
                          weighted.mean(PV1MATH[ST04Q01=="Female"], w=SENWGT_STU[ST04Q01=="Female"], na.rm=T),
                  mtest.stat = t.test(PV1MATH[ST04Q01=="Male"], PV1MATH[ST04Q01=="Female"])$statistic, 
                  mp.value = t.test(PV1MATH[ST04Q01=="Male"], PV1MATH[ST04Q01=="Female"])$p.value,
                  readgap=mean(PV1READ[ST04Q01=="Male"], na.rm=T)-mean(PV1READ[ST04Q01=="Female"], na.rm=T),
                  rtest.stat = t.test(PV1READ[ST04Q01=="Male"], PV1READ[ST04Q01=="Female"])$statistic, 
                  rp.value = t.test(PV1READ[ST04Q01=="Male"], PV1READ[ST04Q01=="Female"])$p.value,
                  sciencegap=mean(PV1SCIE[ST04Q01=="Male"], na.rm=T)-mean(PV1SCIE[ST04Q01=="Female"], na.rm=T),
                  stest.stat = t.test(PV1SCIE[ST04Q01=="Male"], PV1SCIE[ST04Q01=="Female"])$statistic, 
                  sp.value = t.test(PV1SCIE[ST04Q01=="Male"], PV1SCIE[ST04Q01=="Female"])$p.value,                                         
                  minmale=min(PV1MATH[ST04Q01=="Male"], na.rm=T), 
                  minfemale=min(PV1MATH[ST04Q01=="Female"], na.rm=T), 
                  maxmale=max(PV1MATH[ST04Q01=="Male"], na.rm=T), 
                  maxfemale=max(PV1MATH[ST04Q01=="Female"], na.rm=T), 
                  propmale=length(PV1MATH[ST04Q01=="Male"])/length(PV1MATH), 
                  propfemale=length(PV1MATH[ST04Q01=="Female"])/length(PV1MATH))
#qplot(mathgap, wmathgap, data=student2012.sub.summary.gap, xlab="Mean", ylab="Weighted Mean",
#      xlim=c(-30,30), ylim=c(-30,30)) + geom_abline(slope=1) + theme(aspect.ratio=1)
student2012.sub.summary.gap$msig <- ifelse(student2012.sub.summary.gap$mp.value>0.05, "none", TRUE)
student2012.sub.summary.gap$msig[student2012.sub.summary.gap$msig==TRUE&student2012.sub.summary.gap$mtest.stat>0] <- "male"
student2012.sub.summary.gap$msig[student2012.sub.summary.gap$msig==TRUE&student2012.sub.summary.gap$mtest.stat<0] <- "female"
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$mathgap)])
qplot(name, mathgap, data=student2012.sub.summary.gap, size=propmale, color=msig) + 
  xlab("") +
  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  scale_y_continuous("Math Score Gap", breaks=seq(-30, 30, 5)) + 
  scale_size("Prop male") +
   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw() + theme(legend.position="bottom")
```

```{r gendermaps, fig.width=12, fig.height=7, warning=FALSE, message=FALSE, echo=FALSE}
student2012.sub.map <- left_join(student2012.sub.summary.gap, world.polys)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=msig)) +
  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  new_theme_empty + theme(legend.position="none")
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$maxmale)])
#qplot(name, maxmale, data=student2012.sub.summary.gap, color=I("skyblue")) + 
#  xlab("") +
#  geom_point(aes(y=maxfemale), colour="pink") + 
#  scale_y_continuous("High Math") + coord_flip() + theme_bw()
#ggsave("gendermathtop.pdf", width=3.5, height=7)
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$minfemale)])
#qplot(name, minmale, data=student2012.sub.summary.gap, color=I("skyblue")) + 
#  xlab("") +
#  geom_point(aes(y=minfemale), colour="pink") + 
#  scale_y_continuous("Low Math") + coord_flip() + theme_bw()
#ggsave("gendermathbottom.pdf", width=3.5, height=7)
#
```

```{r genderead, fig.width=8, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
student2012.sub.summary.gap$rsig <- ifelse(student2012.sub.summary.gap$rp.value>0.05, "none", TRUE)
student2012.sub.summary.gap$rsig[student2012.sub.summary.gap$rsig==TRUE&student2012.sub.summary.gap$rtest.stat>0] <- "male"
student2012.sub.summary.gap$rsig[student2012.sub.summary.gap$rsig==TRUE&student2012.sub.summary.gap$rtest.stat<0] <- "female"
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$readgap)])
qplot(name, readgap, data=student2012.sub.summary.gap, size=propmale, color=rsig) + 
  xlab("") +
  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  scale_y_continuous("Reading Score Gap") + 
  scale_size("Prop male") +
   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw() + theme(legend.position="bottom")
#ggsave("genderreadgap.pdf", width=5, height=8)
#student2012.sub.summary.gap$ssig <- ifelse(student2012.sub.summary.gap$sp.value>0.05, "none", TRUE)
#student2012.sub.summary.gap$ssig[student2012.sub.summary.gap$ssig==TRUE&student2012.sub.summary.gap$stest.stat>0] <- "male"
#student2012.sub.summary.gap$ssig[student2012.sub.summary.gap$ssig==TRUE&student2012.sub.summary.gap$stest.stat<0] <- "female"
#student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
#      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$sciencegap)])
#qplot(name, sciencegap, data=student2012.sub.summary.gap, size=propmale, color=ssig) + 
#  xlab("") +
#  scale_colour_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
#  scale_y_continuous("Science Score Gap") + 
#  scale_size("Prop male") +
#   geom_hline(yintercept=0, colour="grey80") + coord_flip() + theme_bw()
```

```{r gendermaps2, fig.width=12, fig.height=7, warning=FALSE, message=FALSE, echo=FALSE}
#ggsave("gendermathmap.pdf", width=12, height=7)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=rsig)) +
  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
  new_theme_empty + theme(legend.position="none")
#ggsave("genderreadmap.pdf", width=12, height=7)
#ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
#  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=ssig)) +
#  scale_fill_manual("Significant", values=c("male"="skyblue", "female"="pink", "none"="lightgreen")) +
#  new_theme_empty + theme(legend.position="none")
```
